---
title: "Practical Stats - Chapter 1"
author: "George Khalilieh"
date: "5/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Legal Note:
I'm copying some text verbatim from the book for studying purposes. 
I've already purchased the book and am using this for educational purposes.

## Data Types
There are two basic types of structured data: numeric and categorical. Numeric data comes in two forms:
continuous, such as wind speed or time duration, and discrete, such as the count of the occurrence of an
event. Categorical data takes only a fixed set of values, such as a type of TV screen (plasma, LCD, LED,
etc.) or a state name (Alabama, Alaska, etc.). Binary data is an important special case of categorical data
that takes on only one of two values, such as 0/1, yes/no, or true/false. Another useful type of categorical
data is ordinal data in which the categories are ordered; an example of this is a numerical rating (1, 2, 3,
4, or 5).

## Rectangular Data
The typical frame of reference for an analysis in data science is a rectangular data object, like a
spreadsheet or database table.

### Data frame
Rectangular data (like a spreadsheet) is the basic data structure for statistical and machine learning models.
### Feature
A column in the table is commonly referred to as a feature.
Synonyms
attribute, input, predictor, variable

Rectangular data is essentially a two-dimensional matrix with rows indicating records (cases) and
columns indicating features (variables). The data doesn’t always start in this form: unstructured data (e.g.,
text) must be processed and manipulated so that it can be represented as a set of features in the rectangular
data (see “Elements of Structured Data”). Data in relational databases must be extracted and put into a
single table for most data analysis and modeling tasks.

## Data Frames and Indexes
Traditional database tables have one or more columns designated as an index. This can vastly improve the
efficiency of certain SQLqueries. In Python, with the pandas library, the basic rectangular data structure
is a DataFrame object. By default, an automatic integer index is created for a DataFrame based on the
order of the rows. In pandas, it is also possible to set multilevel/hierarchical indexes to improve the
efficiency of certain operations.
In R, the basic rectangular data structure is a data.frame object. A data.frame also has an implicit
integer index based on the row order. While a custom key can be created through the row.names attribute,
the native R data.frame does not support user-specified or multilevel indexes. To overcome this
deficiency, two new packages are gaining widespread use: data.table and dplyr. Both support
multilevel indexes and offer significant speedups in working with a data.frame.


```{r cars}
summary(cars)
```


```{r pressure, echo=FALSE}
### Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
plot(pressure)
```

